Experimental Analysis of a Spatialised Audio Interface for People with Visual Impairments

### 2019-12-11 paper.tex
Review and submitted!

\articletitle{Experimental Analysis of a Spatialised Audio Interface for People with Visual Impairments}

\articleabstract{
  Sound perception is a fundamental skill for many people with severe sight impairments.
  The research presented in this paper is part of the ActiVis project, the aim of which is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
  This system requires an effective non-visual interface and uses bone-conduction headphones to transmit audio instructions to the user.
  It has been implemented and tested with spatialised audio cues, which convey the direction of a predefined target in 3D space.
  We present an in-depth evaluation of the audio interface with several experiments that involve a large number of participants, both blindfolded and with actual visual impairments, and analyse the pros and cons of our design choices.
  In addition to producing results comparable to the state-of-the-art, we found that Fitts's Law for predicting human movement can be used to form a metric that can be used to improve and refine the quality of the audio interface in future mobile navigation aids. 
}

@@

The ActiVis\footnote{https://lcas.github.io/ActiVis/} project aims to build an indoor mobile navigation aid for people with vision impairments that can guide them to their destination in the last leg of their journey (the so-called `last 10-yard problem').
Improvements to computer vision and machine learning techniques, as well as mobile computing hardware performance increases are exploited to make this system possible.
In particular, techniques from the active vision field are being used to enable a mobile device to gather information on the surrounding environment and use it to generate guidance instructions for a user with limited or no vision.
However, in literature, these techniques are typically limited to directing electro-mechanical servos~\citep{bajcsy2018revisiting} and in this project, we attempt to find out if the same techniques can also be used to direct a user's attention to a certain point. 
In previous work we developed a prototype system that uses active vision and machine learning models to gather information and help a person find objects in an unknown environment, showing that these techniques can indeed successfully be applied to direct humans' attention as well.%~\citep{lock2019active}.

The prototype platform is based on a Google Project Tango\footnote{https://en.wikipedia.org/wiki/Tango\_(platform)} device, pictured in \Cref{fig:tango-headphone}, that embeds a colour camera and provides access to powerful real-time localisation (through IMU measurements and landmark tracking) and image-processing facilities. 
It also provides access to Android's full range of interface tools and I/O options. 
Furthermore, a set of bone-conduction headphones (see \Cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.

@@

Humans are naturally able to determine the 3D position of a sound source and by exploiting this ability, the real-time guidance instructions can be interpreted without posing a significant cognitive load.
A sound source can be spatialised by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle), and intensity (distance).
In our case, only the pan and elevation positions are transmitted to the user to point the camera towards a target object or visual feature.
However, since bone-conduction headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
In previous work, we showed that this audio signal transmission scheme can direct users to a target position with a degree of accuracy comparable to fully spatialised signals used with expensive closed-cup headphones and we expand upon this initial investigation with a larger dataset and look at if, and how, changing the behaviour of the pitch affects target acquisition performance in terms of time and angular error.% ~\citep{lock2019bone,macdonald2006spatial}
The participants' hearing characteristics are also measured to determine if there are limitations to how well they can determine audio pitch or direction.

@@

Simple audio tones and vocal feedback signals have been used in conjunction with cameras and object detectors to communicate their relative positions to the user with a reasonable level of accuracy~\citep{schauerte2012assistive,tian2013computer,fiannaca2014headlock,vazquez2012helping}.
Following this, researchers have investigated using audio signals that are spatialised with a head-related transfer function (HRTF), simulating a sound source located at some arbitrary 3D position~\citep{geronazzo2016interactive,wilson2007swan,katz2010navig,blum2013spatialized}.
The authors generally report favourable results for this guidance approach when used with standard over-ear headphones or speakers. 
However, other authors have found that the choice of audio transmission medium can have a significant effect on performance.
Cheaper headphones and bone-conduction headphones generally report unfavourable results when compared to over-ear or more expensive alternatives~\citep{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}. 
This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conduction pathway~\citep{stanley2006lateralization}.
~\cite{durette2008visuo} transmit a target's elevation angle by adjusting the signal's pitch and report favourable results. 
We extended on the latter's work and investigated spatial audio for guidance mode closely in an initial study and found that it is indeed able to adequately guide a user towards a target area.%~\citep{lock2019bone}.
We expand upon this preliminary study with a more thorough investigation into the participants' performance and how changes to the audio signal's behaviour can impact that performance.

@@

Existing electronic navigation aids have typically struggled to gain market traction and replace the traditional walking cane as the standard assistive tool for people with visual impairments.
Current technological limitations include prohibitive costs, bulky hardware requirements and non-user-friendly interfaces~\citep{golledge2004stated,yusif2016older,arditi2013user}.
To address these issues, we implemented a handheld mobile system using a Google Tango device that is able to localise itself in real-time.
%To address these issues, we implemented a handheld mobile system that is based on a concept proposed by~\cite{lock2019active} and tested by~\cite{lock2019bone} using a Google Tango device that is able to localise itself in real-time.
This system has the benefit of minimal hardware requirements and a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.

@@

Humans localise a sound source in three dimensions by considering cues recorded in one ear (monaural cues) and comparing cues received at both ears (binaural cues)~\citep{blauert1997spatial,blauert1969sound}.
The binaural cues include inter-aural time and level differences (ITD and ILD respectively) that help to determine a source's location on the horizontal plane.
Monaural cues are taken from the interaction of the sound with the human anatomy, e.g.\ head, shoulders, outer ear, before it enters the ear canal.
When the modified audio signal enters the inner ear canal, the human brain is able to analyse the frequency response and accurately determine the position of the sound source on the median plane. 
The distance to the source is simply derived as the intensity, or volume, of the source, i.e.\ a louder sound would appear closer to the user than a softer one.

@@

The guidance information is presented to the user in terms of pan and elevation angles, indicating the angular adjustments required to point the device camera at the target location, as shown in \Cref{fig:cam-coords}.
Spatialised audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution~\citep{klatzky2006cognitive}.
However, given the previously discussed limitations of bone-conduction and spatialised audio, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
The pan angle, instead, can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conduction headphones~\citep{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.

@@

The human audition system uses binaural comparison cues, such as ITD and ILD, to localise a sound source on the horizontal plane~\citep{blauert1969sound}.
The ITD is the perceived time delay between the signal reaching both ears, while the ILD is the perceived volume difference in the signal.
For example, a sound that comes from the individual's right will hit the right ear first with a slightly higher volume compared to the left ear.

In this work, a pure sinusoidal wave was used. 
People typically have trouble localising a pure tone without a sufficiently rich spectral signature.
However, the ITD and ILD are independent of the tone's spectral make-up, while the elevation angle is given through a different mechanism. Therefore, a pure sine wave is suitable to convey the target's pan angle.
To transform and spatialise the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\citep{hiebert2005openal}, which uses the person and targets' positions as input, and outputs a transformed audio signal.

@@

We performed a set of experiments with the audio interface to determine how effective it is at directing a user to adjust the pan and elevation angles of a camera for point it to a target.
Furthermore, we also carried out a set of pre-screening experiments to determine each participant's hearing characteristics in order to determine their perception limits in the respective audio dimensions.
The participants were given time before the experiments commenced to familiarise themselves with the device and the tones it emits, as well as what the `on-target' tone sounds like.
We also tested the system with a group of severely sight impaired participants and compared their data to the blindfolded dataset.
The results from the experiments we performed allow us to better understand how the users respond to different settings for the spatial audio feedback stimulus and use them to improve and optimise the behaviour of the feedback modes in our portable navigation aid.

@@

\subsection{Participant Characterisation}\label{sec:participant_characterisation}

A preliminary set of experiments were conducted to characterise the participants' hearing characteristics.
The measured characteristics were each participant's audio localisation ability on the lateral plane, as well as the participants' ability to discriminate between tones with different frequencies. 
These results will be used to provide context to the following target search experiment and provide additional insight on any possible biases or limitations.

\subsubsection{Sound Localisation}\label{sec:sound_localisation}

In this experiment, we evaluated a participant's ability to determine the lateral direction a sound is coming from.
To do this, we played a continuous \SI{512}{\hertz} sinusoidal tone to the participant through the headphones and applied an HRTF to spatialise and place its source to the participant's left or right.
The participant then had to select the direction the sound came from.
The longer the experiment lasts and the more correct guesses the participant makes, the closer the source moves to the centre-front of the participant, making it increasingly harder to localise.

For this progressive increase in difficulty, a ``2-up, 1-down'' step process is used~\citep{wetherill1965sequential,levitt1971transformed}, meaning that for every two correct answers, the distance to the centre halves.
Conversely, the task becomes easier for each incorrect answer by doubling the sound source's distance from the centre.
We also use two different step sequences, one starting at a large angular distance (\SI{45}{\degree}) from the user and the other at the minimum distance (approximately \SI{1}{\degree}), giving an `easy' and a `hard' progression respectively.
The terminating condition for the experiment is when the two sequences converge to within two intervals of one another for three consecutive guesses.
For example, the experiment will terminate when one sequence is set to \SI{11.25}{\degree} and the other is between \SI{2.8}{\degree} or \SI{45}{\degree} for three consecutive guesses.
This gives a distance band where the participant is capable of localising the sound source.
Each participant performed this experiment three times.

@@

The first tone is randomly generated, while the second tone is generated by adding or subtracting some value from the first one.
The tone difference depends on how well the participant can tell the tones apart.
Like the sound localisation experiment, a ``2-up, 1-down'' step process is used: for every two consecutive correct answers, the pitch difference between the tones is halved, while it is doubled for every incorrect answer.
Two-step sequences are again used here, one starting with a large pitch difference ($f_h=2^9=$ \SI{512}{\hertz}) between the tones and the other with a small difference ($f_l=2^1=$ \SI{2}{\hertz}).
The termination condition is when the two-step sequences are within one octave of each other (i.e.\ $\log_2\frac{f_h}{f_l}=2$) for three consecutive answers.
For example, the experiment will terminate when one sequence is set to \SI{64}{\hertz} and the other is between \SI{32}{\hertz} or \SI{128}{\hertz} for three consecutive guesses.
Pitch differences are measured in semitones, which can be obtained with

@@

To test the interface's effectiveness at guiding the user in a pointing task, a set of experiments were conducted to capture the difference between the targets' actual direction and the directions the participants' perceived them to be.
The participants were given a Tango device running an app that implements the experimental setup in~\Cref{fig:pipeline}. The app generates a set of virtual targets and presents them to each participant through the audio interface, one at a time. 
The targets are generated at a constant distance from the participant and their pan and elevation angles are uniformly generated across the four quadrants of the pan-elevation plane to avoid clustering.
Each target's angular position is adjusted and communicated in real-time as the participant points the device around. 
When the participants were confident that the device was on-target, i.e.~hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
The targets' positions are all set relative to the device's coordinate system, which is tracked using the Tango hardware and localisation API.\
A total of 28 targets were generated per participant.

As part of this experiment, we wanted to see how changing the gradient of the pitch function, visualised in~\Cref{fig:pitch-gain}, affects target acquisition performance, e.g.\ does a steeper pitch gain as a function of the elevation angle improve accuracy or decrease the search time?
Pitch limits of one, two and three octaves above and below the neutral tone were then set for the so-called \textit{lo}, \textit{med} and \textit{hi} pitch gradient settings respectively, giving pitch intervals of

@@

Here we briefly summarise the equations and quantities involved in our metric.
Fitts's Law is given by

@@

\noindent
Here $\theta$ is the angular distance between subsequent target centres and $w_e$ is the targets' effective angular width~\citep{welford1968fundamentals}, given by

@@

Two groups of participants were recruited for the experiments on a volunteer basis. 
Group~\textit{G1} consisted of 42 young adults with normal eyesight who were blindfolded for the experiments, and group \textit{G2} contained 10 people with severe visual impairments (18--55 years old; 36 female, 16 male). 
None of the participants had any hearing or other disabilities that could have influenced their performance in the experiments.
Each participant performed three sets of experiments each, with the two characterisation experiments in \Cref{sec:participant_characterisation} preceding the final target-search experiment in \Cref{sec:target_search}. 
Both groups were given some time before the target-search experiment to familiarise themselves with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
Furthermore, to minimise any potential speed/accuracy biases, we asked the participants to focus on finding the targets without worrying about the time it took to complete the task.

@@

\subsection{Characterisation of Sound Localisation}

\Cref{fig:sound-localisation} shows the results captured from the sound localisation experiment where the participants had to select the direction (left or right) that the tone was being played from. 
%
It can be seen that the vast majority of guesses for both groups were correct.
For Group~\textit{G1}, most of the errors were made at the minimum distance from the centre, i.e.\ the most difficult to guess correctly, which is the expected behaviour.
This indicates that the participants in \textit{G1} consistently progressed through the distance intervals and we can therefore conclude they had little difficulty determining sound direction.

@@

These results show that both participant groups are capable of determining a sound source's location with a reasonable level of consistency and accuracy.
These results are in line previous literature, confirming that humans are very adept at localising a sound source, particularly in the pan dimension.

\subsection{Characterisation of Pitch Discrimination}

@@

The results from the target search experiment shown on the 2D histograms in \Cref{fig:target-errors}, where the angular errors in the pan and elevation dimensions are plotted against each other. 
A set of box-plots of the angle errors are also given in \Cref{fig:target-boxplot-error} for each audio setting.
The results are summarised in \Cref{tab:target-results}.

@@

Comparing the distributions for each setting between the two groups with the Kruskal-Wallis test for non-parametric data, we see that the differences between the distributions for all three settings are not significantly different, for both groups and both pan and elevation (the $p$-values are summarised in \Cref{tab:inter-group-results}).
This results confirm that the performance of the blindfolded participants and those with severe sight impairments are statistically similar, and that groups from a different population can reasonably be expected to produce similar errors, under similar experimental conditions. 
Consequently, we can conclude that the \textit{hi} setting, which generates the significantly smallest elevation error, is the best audio pitch level to guide a user in a pointing task, and that the pan error is completely independent of such setting choice.

@@

These results allows us to calculate and plot in \Cref{fig:fitts-ips} an index of performance, as given by \Cref{eq:fitts-performance}, for each audio setting.
The results are summarised in \Cref{tab:fitts-results}.
For Group~\textit{G1}, there is a fairly consistent level of performance between the three settings, with \textit{lo} producing the highest indices of performance overall (i.e.\ the participants found the targets with the smallest error and in the least amount of time).
This is supported by the results from the Friedman test, showing that there is a significant difference in performance between the settings ($p < 0.001$), as well as post-hoc Wilcoxon tests with Holm-Bonferroni corrections, which show that the \textit{lo} setting is significantly different to the \textit{med} and \textit{hi} settings ($p_{lo-med} < 0.001,~p_{lo-hi} < 0.001$). The \textit{med} and \textit{hi}, instead, are not significantly different from each other ($p_{med-hi} = 0.85$).
The results for Group~\textit{G2} show generally lower and inconsistent indices of performance for each setting, which is expected given the increased times to target observed in \Cref{fig:fitts-results}.
Again, from the Friedman test ($p < 0.001$), the \textit{lo} setting produces the highest performance by a large margin, compared to the \textit{med} setting, with the Wilcoxon test and Holm-Bonferronni corrections ($p_{lo-med}=0.01$), followed by the \textit{hi} and \textit{med} settings' results, respectively.
This seems to indicate that, for both groups, the \textit{lo} setting produces the highest level of performance, followed by the \textit{hi} setting.

@@

The results for the accuracy and time performance with the proposed audio interface shows an interesting contrast.
That is, the \textit{hi} pitch setting produces the lowest target acquisition error, followed by the \textit{med} and \textit{lo} settings, respectively. 
However, this trend is almost wholly reversed in the time-to-target results from the Fitts model, where the \textit{lo} setting gives the highest level of performance, followed by the \textit{hi} and \textit{med} settings respectively. 
Since Fitts's model takes the angular error into account, one might reasonably expect that the results for both experiments follow a similar trend. 
However, the Fitts model does not account for changing stimuli and different strategies.
We therefore hypothesise that the reason for this divergence in performance is due to the increased resolution of the \textit{hi} setting, which allows for finer adjustments of the device's orientation, getting it closer to the correct target, but at the cost of a higher average time-to-target. 
This seems to indicate a speed/accuracy trade-off in finding the targets.
With the Fitts model discussed here, we can modify the interface to prioritise different metrics and produce the correct output.

@@

In this paper we investigated the use of spatialised audio interface with varying pitch to guide a visually impaired user in a target pointing task.
We found that the blindfolded participants and those with severe sight impairments performed similarly in localising sound sources and differentiating between tones. 
We also found that both groups were able to find a randomly distributed set of virtual targets with similar levels of accuracy.
However, the blindfolded group outperformed the one with severe sight impairments in terms of time-to-target. 
We further tested different pitch settings and found that the user performance in the pan dimension, based on spatialised cues, is independent of such settings.
Moreover, we noticed a speed/accuracy trade-off between the settings, where a higher pitch setting produces a smaller angular error, but at the cost of reducing the time performance (i.e.\ more time to reach the target). 
These results, together with an analysis done with Fitts's Law that confirms its applicability to this type of audio interface, provide a useful baseline to improve and refine the latter in future applications, prioritising speed or accuracy to produce the desired output.

>>>>

\articletitle{Experimental Analysis of a Spatialized Audio Interface for People with Visual Impairments}

\articleabstract{
  Sound perception is a fundamental skill for many people with severe sight impairments.
  The research presented in this paper is part of an ongoing project, the aim of which is to build a mobile guidance aid to help people with limited vision find objects in an unknown environment.
  This system requires an effective non-visual interface and uses bone-conduction headphones to transmit audio instructions to the user.
  It has been implemented and tested with spatialized audio cues, which convey the direction of a predefined target in 3D space.
  We present an in-depth evaluation of the audio interface with several experiments that involve a large number of participants, both blindfolded and with actual visual impairments, and analyze the pros and cons of our design choices.
  In addition to producing results comparable to the state-of-the-art, we found that Fitts's Law for predicting human movement can be used to form a metric that can be used to improve and refine the quality of the audio interface in future mobile navigation aids. 
}

@@

This project aims to build an indoor mobile navigation aid for people with vision impairments that can guide them to their destination in the last leg of their journey (the so-called `last 10-yard problem').
Improvements to computer vision and machine learning techniques, as well as mobile computing hardware performance increases are exploited to make this system possible.
In particular, techniques from the active vision field are used to enable a mobile device to gather information on the surrounding environment and use it to generate guidance instructions for a user with limited or no vision.
However, in literature, these techniques are typically limited to directing electro-mechanical servos~\citep{bajcsy2018revisiting} and in this project, we attempt to find out if the same techniques can also be used to direct a user's attention to a certain point. 
In previous work we developed a prototype system that uses active vision and machine learning models to gather information and help a person find objects in an unknown environment, showing that these techniques can indeed successfully be applied to direct humans' attention as well.%~\citep{lock2019active}.

The prototype platform is based on a Google Project Tango\footnote{https://en.wikipedia.org/wiki/Tango\_(platform)} device, pictured in \Cref{fig:tango-headphone}, that embeds a color camera and provides access to powerful real-time localization (through IMU measurements and landmark tracking) and image-processing facilities. 
It also provides access to Android's full range of interface tools and I/O options. 
Furthermore, a set of bone-conduction headphones (see \Cref{fig:tango-headphone}), which are placed on a user's cheekbones instead of their ears and do not interfere with normal hearing, are used to transmit the audio signals to the user.

@@

Humans are naturally able to determine the 3D position of a sound source and by exploiting this ability, the real-time guidance instructions can be interpreted without posing a significant cognitive load.
A sound source can be spatialized by adjusting a tone's spectral make-up (elevation angle), time delay and level difference (pan angle), and intensity (distance).
In our case, only the pan and elevation positions are transmitted to the user to point the camera towards a target object or visual feature.
However, since bone-conduction headphones bypass the outer ear structure, their spectral signature cannot properly be interpreted and we therefore convey the target's elevation angle by adjusting the tone's pitch.
In previous work, we showed that this audio signal transmission scheme can direct users to a target position with a degree of accuracy comparable to fully spatialized signals used with expensive closed-cup headphones and we expand upon this initial investigation with a larger dataset and look at if, and how, changing the behavior of the pitch affects target acquisition performance in terms of time and angular error.% ~\citep{lock2019bone,macdonald2006spatial}
The participants' hearing characteristics are also measured to determine if there are limitations to how well they can determine audio pitch or direction.

@@

Simple audio tones and vocal feedback signals have been used in conjunction with cameras and object detectors to communicate their relative positions to the user with a reasonable level of accuracy~\citep{schauerte2012assistive,tian2013computer,fiannaca2014headlock,vazquez2012helping}.
Following this, researchers have investigated using audio signals that are spatialized with a head-related transfer function (HRTF), simulating a sound source located at some arbitrary 3D position~\citep{geronazzo2016interactive,wilson2007swan,katz2010navig,blum2013spatialized}.
The authors generally report favorable results for this guidance approach when used with standard over-ear headphones or speakers. 
However, other authors have found that the choice of audio transmission medium can have a significant effect on performance.
Cheaper headphones and bone-conduction headphones generally report unfavorable results when compared to over-ear or more expensive alternatives~\citep{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}. 
This seems to be limited to the elevation dimension, however, and can be improved with HRTFs adjusted for the bone-conduction pathway~\citep{stanley2006lateralization}.
~\cite{durette2008visuo} transmit a target's elevation angle by adjusting the signal's pitch and report favorable results. 
We extended on the latter's work and investigated spatial audio for guidance mode closely in an initial study and found that it is indeed able to adequately guide a user towards a target area.%~\citep{lock2019bone}.
We expand upon this preliminary study with a more thorough investigation into the participants' performance and how changes to the audio signal's behavior can impact that performance.

@@

Existing electronic navigation aids have typically struggled to gain market traction and replace the traditional walking cane as the standard assistive tool for people with visual impairments.
Current technological limitations include prohibitive costs, bulky hardware requirements and non-user-friendly interfaces~\citep{golledge2004stated,yusif2016older,arditi2013user}.
To address these issues, we implemented a handheld mobile system using a Google Tango device that is able to localize itself in real-time.
%To address these issues, we implemented a handheld mobile system that is based on a concept proposed by~\cite{lock2019active} and tested by~\cite{lock2019bone} using a Google Tango device that is able to localise itself in real-time.
This system has the benefit of minimal hardware requirements and a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.

@@

Humans localize a sound source in three dimensions by considering cues recorded in one ear (monaural cues) and comparing cues received at both ears (binaural cues)~\citep{blauert1997spatial,blauert1969sound}.
The binaural cues include inter-aural time and level differences (ITD and ILD respectively) that help to determine a source's location on the horizontal plane.
Monaural cues are taken from the interaction of the sound with the human anatomy, e.g.\ head, shoulders, outer ear, before it enters the ear canal.
When the modified audio signal enters the inner ear canal, the human brain is able to analyze the frequency response and accurately determine the position of the sound source on the median plane. 
The distance to the source is simply derived as the intensity, or volume, of the source, i.e.\ a louder sound would appear closer to the user than a softer one.

@@

The guidance information is presented to the user in terms of pan and elevation angles, indicating the angular adjustments required to point the device camera at the target location, as shown in \Cref{fig:cam-coords}.
Spatialized audio signals are well-suited to the task, displaying similar levels of performance to vocal feedback, but with less cognitive load and higher resolution~\citep{klatzky2006cognitive}.
However, given the previously discussed limitations of bone-conduction and spatialized audio, we propose a simple linear adjustment to the signal's pitch as a function of the elevation angle. 
The pan angle, instead, can be conveyed by transforming the audio signal with an HRTF, and indeed it has been found that this dimension is unaffected by using bone-conduction headphones~\citep{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization}.

@@

The human audition system uses binaural comparison cues, such as ITD and ILD, to localize a sound source on the horizontal plane~\citep{blauert1969sound}.
The ITD is the perceived time delay between the signal reaching both ears, while the ILD is the perceived volume difference in the signal.
For example, a sound that comes from the individual's right will hit the right ear first with a slightly higher volume compared to the left ear.

In this work, a pure sinusoidal wave was used. 
People typically have trouble localizing a pure tone without a sufficiently rich spectral signature.
However, the ITD and ILD are independent of the tone's spectral make-up, while the elevation angle is given through a different mechanism. Therefore, a pure sine wave is suitable to convey the target's pan angle.
To transform and spatialize the audio signal, we used OpenAL's default HRTF, based on the MIT's KEMAR dataset~\citep{hiebert2005openal}, which uses the person and targets' positions as input, and outputs a transformed audio signal.

@@

We performed a set of experiments with the audio interface to determine how effective it is at directing a user to adjust the pan and elevation angles of a camera for point it to a target.
Furthermore, we also carried out a set of pre-screening experiments to determine each participant's hearing characteristics in order to determine their perception limits in the respective audio dimensions.
The participants were given time before the experiments commenced to familiarize themselves with the device and the tones it emits, as well as what the `on-target' tone sounds like.
We also tested the system with a group of participants with severe sight impairments and compared their data to the blindfolded participant dataset.
The results from the experiments we performed allow us to better understand how the users respond to different settings for the spatial audio feedback stimulus and use them to improve and optimize the behavior of the feedback modes in our portable navigation aid.

@@

\subsection{Participant Characterization}\label{sec:participant_characterisation}

A preliminary set of experiments were conducted to characterize the participants' hearing characteristics.
The measured characteristics were each participant's audio localization ability on the lateral plane, as well as the participants' ability to discriminate between tones with different frequencies. 
These results will be used to provide context to the following target search experiment and provide additional insight on any possible biases or limitations.

\subsubsection{Sound Localization}\label{sec:sound_localisation}

In this experiment, we evaluated a participant's ability to determine the lateral direction a sound is coming from.
To do this, we played a continuous \SI{512}{\hertz} sinusoidal tone to the participant through the headphones and applied an HRTF to spatialize and place its source to the participant's left or right.
The participant then had to select the direction the sound came from.
The longer the experiment lasts and the more correct guesses the participant makes, the closer the source moves to the centre-front of the participant, making it increasingly harder to localize.

For this progressive increase in difficulty, a ``2-up, 1-down'' step process is used~\citep{wetherill1965sequential,levitt1971transformed}, meaning that for every two correct answers, the distance to the centre halves.
Conversely, the task becomes easier for each incorrect answer by doubling the sound source's distance from the centre.
We also use two different step sequences, one starting at a large angular distance (\SI{45}{\degree}) from the user and the other at the minimum distance (approximately \SI{1}{\degree}), giving an `easy' and a `hard' progression respectively.
The terminating condition for the experiment is when the two sequences converge to within two intervals of one another for three consecutive guesses.
For example, the experiment will terminate when one sequence is set to \SI{11.25}{\degree} and the other is between \SI{2.8}{\degree} or \SI{45}{\degree} for three consecutive guesses.
This gives a distance band where the participant is capable of localizing the sound source.
Each participant performed this experiment three times.

@@

The first tone is randomly generated, while the second tone is generated by adding or subtracting some value from the first one.
The tone difference depends on how well the participant can tell the tones apart.
Like the sound localization experiment, a ``2-up, 1-down'' step process is used: for every two consecutive correct answers, the pitch difference between the tones is halved, while it is doubled for every incorrect answer.
Two-step sequences are again used here, one starting with a large pitch difference ($f_h=2^9=$ \SI{512}{\hertz}) between the tones and the other with a small difference ($f_l=2^1=$ \SI{2}{\hertz}).
The termination condition is when the two-step sequences are within one octave of each other (i.e.\ $\log_2\frac{f_h}{f_l}=2$) for three consecutive answers.
For example, the experiment will terminate when one sequence is set to \SI{64}{\hertz} and the other is between \SI{32}{\hertz} or \SI{128}{\hertz} for three consecutive guesses.
Pitch differences are measured in semitones, which can be obtained with

@@

To test the interface's effectiveness at guiding the user in a pointing task, a set of experiments were conducted to capture the difference between the targets' actual direction and the directions the participants' perceived them to be.
The participants were given a Tango device running an app that implements the experimental setup in~\Cref{fig:pipeline}. The app generates a set of virtual targets and presents them to each participant through the audio interface, one at a time. 
The targets are generated at a constant distance from the participant and their pan and elevation angles are uniformly generated across the four quadrants of the pan-elevation plane to avoid clustering.
Each target's angular position is adjusted and communicated in real-time as the participant points the device around. 
When the participants were confident that the device was on-target, i.e.~hearing the audio front-on at \SI{512}{\hertz}, they tapped the screen, marking the location and generating the next target.
The targets' positions are all set relative to the device's coordinate system, which is tracked using the Tango hardware and localization API.\
A total of 28 targets were generated per participant.

As part of this experiment, we wanted to see how changing the gradient of the pitch function, visualized in~\Cref{fig:pitch-gain}, affects target acquisition performance, e.g.\ does a steeper pitch gain as a function of the elevation angle improve accuracy or decrease the search time?
Pitch limits of one, two and three octaves above and below the neutral tone were then set for the so-called \textit{lo}, \textit{med} and \textit{hi} pitch gradient settings respectively, giving pitch intervals of

@@

Here we briefly summarize the equations and quantities involved in our metric.
Fitts's Law is given by

@@

\noindent
Here $\theta$ is the angular distance between subsequent target centers and $w_e$ is the targets' effective angular width~\citep{welford1968fundamentals}, given by

@@

Two groups of participants were recruited for the experiments on a volunteer basis. 
Group~\textit{G1} consisted of 42 young adults with normal eyesight who were blindfolded for the experiments, and group \textit{G2} contained 10 people with severe visual impairments (18--55 years old; 36 female, 16 male). 
None of the participants had any hearing or other disabilities that could have influenced their performance in the experiments.
Each participant performed three sets of experiments each, with the two characterization experiments in \Cref{sec:participant_characterisation} preceding the final target-search experiment in \Cref{sec:target_search}. 
Both groups were given some time before the target-search experiment to familiarize themselves with the system, the audio signal's behavior and the \SI{512}{\hertz} on-level tone. 
Furthermore, to minimize any potential speed/accuracy biases, we asked the participants to focus on finding the targets without worrying about the time it took to complete the task.

@@

\subsection{Characterization of Sound Localization}

\Cref{fig:sound-localisation} shows the results captured from the sound localization experiment where the participants had to select the direction (left or right) that the tone was being played from. 
It can be seen that the vast majority of guesses for both groups were correct.
For Group~\textit{G1}, most of the errors were made at the minimum distance from the centre, i.e.\ the most difficult to guess correctly, which is the expected behavior.
This indicates that the participants in \textit{G1} consistently progressed through the distance intervals and we can therefore conclude they had little difficulty determining sound direction.

@@

These results show that both participant groups are capable of determining a sound source's location with a reasonable level of consistency and accuracy.
These results are in line previous literature, confirming that humans are very adept at localizing a sound source, particularly in the pan dimension.

\subsection{Characterization of Pitch Discrimination}

@@

The results from the target search experiment shown on the 2D histograms in \Cref{fig:target-errors}, where the angular errors in the pan and elevation dimensions are plotted against each other. 
A set of box-plots of the angle errors are also given in \Cref{fig:target-boxplot-error} for each audio setting.
The results are summarized in \Cref{tab:target-results}.

@@

Comparing the distributions for each setting between the two groups with the Kruskal-Wallis test for non-parametric data, we see that the differences between the distributions for all three settings are not significantly different, for both groups and both pan and elevation (the $p$-values are summarized in \Cref{tab:inter-group-results}).
This results confirm that the performance of the blindfolded participants and those with severe sight impairments are statistically similar, and that groups from a different population can reasonably be expected to produce similar errors, under similar experimental conditions. 
Consequently, we can conclude that the \textit{hi} setting, which generates the significantly smallest elevation error, is the best audio pitch level to guide a user in a pointing task, and that the pan error is completely independent of such setting choice.

@@

These results allows us to calculate and plot in \Cref{fig:fitts-ips} an index of performance, as given by \Cref{eq:fitts-performance}, for each audio setting.
The results are summarized in \Cref{tab:fitts-results}.
For Group~\textit{G1}, there is a fairly consistent level of performance between the three settings, with \textit{lo} producing the highest indices of performance overall (i.e.\ the participants found the targets with the smallest error and in the least amount of time).
This is supported by the results from the Friedman test, showing that there is a significant difference in performance between the settings ($p < 0.001$), as well as post-hoc Wilcoxon tests with Holm-Bonferroni corrections, which show that the \textit{lo} setting is significantly different to the \textit{med} and \textit{hi} settings ($p_{lo-med} < 0.001,~p_{lo-hi} < 0.001$). The \textit{med} and \textit{hi}, instead, are not significantly different from each other ($p_{med-hi} = 0.85$).
The results for Group~\textit{G2} show generally lower and inconsistent indices of performance for each setting, which is expected given the increased times to target observed in \Cref{fig:fitts-results}.
Again, from the Friedman test ($p < 0.001$), the \textit{lo} setting produces the highest performance by a large margin, compared to the \textit{med} setting, with the Wilcoxon test and Holm-Bonferronni corrections ($p_{lo-med}=0.01$), followed by the \textit{hi} and \textit{med} settings' results, respectively.
This seems to indicate that, for both groups, the \textit{lo} setting produces the highest level of performance, followed by the \textit{hi} setting.

@@

The results for the accuracy and time performance with the proposed audio interface shows an interesting contrast.
That is, the \textit{hi} pitch setting produces the lowest target acquisition error, followed by the \textit{med} and \textit{lo} settings, respectively. 
However, this trend is almost wholly reversed in the time-to-target results from the Fitts model, where the \textit{lo} setting gives the highest level of performance, followed by the \textit{hi} and \textit{med} settings respectively. 
Since Fitts's model takes the angular error into account, one might reasonably expect that the results for both experiments follow a similar trend. 
However, the Fitts model does not account for changing stimuli and different strategies.
We therefore hypothesize that the reason for this divergence in performance is due to the increased resolution of the \textit{hi} setting, which allows for finer adjustments of the device's orientation, getting it closer to the correct target, but at the cost of a higher average time-to-target. 
This seems to indicate a speed/accuracy trade-off in finding the targets.
With the Fitts model discussed here, we can modify the interface to prioritize different metrics and produce the correct output.

@@

In this paper we investigated the use of spatialized audio interface with varying pitch to guide a user with visual impairments in a target pointing task.
We found that the blindfolded participants and those with severe sight impairments performed similarly in localizing sound sources and differentiating between tones. 
We also found that both groups were able to find a randomly distributed set of virtual targets with similar levels of accuracy.
However, the blindfolded group outperformed the one with severe sight impairments in terms of time-to-target. 
We further tested different pitch settings and found that the user performance in the pan dimension, based on spatialized cues, is independent of such settings.
Moreover, we noticed a speed/accuracy trade-off between the settings, where a higher pitch setting produces a smaller angular error, but at the cost of reducing the time performance (i.e.\ more time to reach the target). 
These results, together with an analysis done with Fitts's Law that confirms its applicability to this type of audio interface, provide a useful baseline to improve and refine the latter in future applications, prioritising speed or accuracy to produce the desired output.
